---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data","learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df #just copying the dataframe
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

There are multiple means of removin

```{r}
df_No_NA <- df %>%
  drop_na()
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

We would run a statistical test (ANOVA) to determine if there is a significant difference in growth rates between treatment levels. Ooops. Didn't read below)

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect? There are 2 DF for the treatment because there are only 3 treatments in total. For each one treatment fixed, the other two can vary. 3-2=1

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this:

```{r}
sum(!is.na(df$Growth_rate_mm))
```

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

49 - 3 = 46 = degrees of freedom for the residuals.

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control? Adding a one dose of fertilizer does not change the growth rate significantly different from the control (p=0.0812)

Does adding 2 doses of fertilizer change the growth rate significantly from the Control? Adding a two doses of fertilizer changes the growth rate significantly from the control, the fertilizer treatment having higher growth rates (p=0.0108)

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(x=Treatment, y=Growth_rate_mm)) +
  geom_boxplot() +
  theme_bw()
```

What do you think?

#### Create a model:

```{r}
grow.mod.no.NA <- lm(Growth_rate_mm ~ Treatment, data=df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA?

They look pretty similar to me. The largest difference is in the Residuals vs Leverage plot in which the leverage in the second ANOVA is less than the first.

#### Interpret results:

```{r}
anova(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

There are 3 treatment levels, 3-1=2.

Why are there 33 DF for the Residuals?

There are 36 observations void of NAs. 36-3=33.

How would you write a results statement?

We found no significant difference in the growth rates between different treatments of fertilizer (df =2, F=2.7906, p=0.0759)

```{r}
summary(grow.mod.no.NA)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control? No. There is no significant difference between the control and the one dose of fertilizer treatment. (p=0.1139)

Does adding 2 doses of fertilizer change the growth rate significantly from the Control? There is a significant difference in growth rates between the control and the two doses of fertilizer treatment, the latter having quicker growth rates. (p=0.0325)

We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(df_All_NA, aes(x=Growth_rate_mm, y=Dried_mass_g)) + 
  geom_point() + 
  theme_bw()
```

#### Create model

```{r}
grow_mass_model <- lm(Dried_mass_g ~ Growth_rate_mm, data=df_All_NA)
```

#### Check assumptions

```{r}
autoplot(grow_mass_model)
```

Don't worry if assumptions don't look perfect - focus on normality of the data.

Everything else looks funky but its normal looking!

#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(grow_mass_model)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals?

There is only one "treatment level" because both variables are numeric. There are 36 entries (points on the graph) and then subtract 2 (for variables).

```{r}
summary(grow_mass_model) 
```

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(df_No_NA, aes(x=Growth_rate_mm, y=Dried_mass_g)) + 
  geom_point() +
  theme_bw()
```

#### Create model

```{r}
grow_mass_model_no_na <- lm(Dried_mass_g ~ Growth_rate_mm, data = df_No_NA)
```

#### Check assumptions

```{r}
autoplot(grow_mass_model_no_na)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis? They look pretty much the same to me.

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(grow_mass_model_no_na)
summary(grow_mass_model_no_na)
```

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why?

    There is only one degree of freedom because there are only two variables, each with only one treatment level.

2.  How many degrees of freedom are there for Residuals? Why?

    There are 34 degrees of freedom for the residuals because there are 36 observation points.

3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis?

    The degrees of freedom does not differ between the two different models. This hints to me that NA values are automatically eliminated in linear regressions. this makes sense because every point on the graph is a value of both variables.
