---
title: "Homework 23"
author: "Evelyn Albrecht"
format: html
editor: visual
---

# 5.4 Introducing... linear models

general linear models include many types of models including regression, multiple regression, ANOVA, and ANCOVA.

assumption of the normally distributed.

We are continuing with out pre-established workflow (plot model check assumptions interpret plot again)

# 5.5 Simple linear regression

example of whether plant growth rates vary with soil moisture content. We predict that higher moisture will allow for higher growth rates.

We have response and explanatory variables. Both are continuous.

## 5.5.1 Getting and plotting the data

set up

```{r}
rm(list=ls())
library(tidyverse)
library("here")
```

Read in the data

```{r}
plant_growth <- read.csv(here("Data","plant.growth.rate.csv"))
glimpse(plant_growth)
```

We can see the data expected.

Make a scatterplot

```{r}
ggplot(plant_growth, aes(x=soil.moisture.content, y=plant.growth.rate))+ 
  geom_point(size = 3, color= "orange")
```

## 5.5.2 Interpreting the figure: biological insight 

A couple of things that we should notice about this graph

-   positive slope

-   pre-empt the statistical analysis

    -   guesstimate the slope and intercept

we should generally try to examine the data before we do the analysis itself. We can even try to figure out degrees freedom.

## 5.5.3 Making a simple linear regression happen 

the function lm() is used to fit a linear model. It is similar to xtabs and t.test in that it needs formulas and data.

```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, data = plant_growth)
```

gonna quote book verbatim here " this reads, 'Fit linear model, where we hypothesize that plant growth is a function of soil moisture content, using the variables from the plant_growth dataframe'

## 5.5.4 Assumptions first 

Now we want to see whether our predictions based on the on the initial plot was correct.

We should first check out assumptions of our model.

new terms!! ggfortify and its autoplot() function.

autoplot uses what we made in the lm() and creates four

I am not sure if ggfortify is part of tidyverse or not

```{r}
install.packages("ggfortify")
library(ggfortify)
autoplot(model_pgr, smooth.colour = NA)
```

"What, pray tell, do these plots mean?"

residuals - errors around a fitted line

the four plots

-    top left -systemic part - indicates whether things have gone wrong (hump-shape valleys will be present)

-    top right - evaluates the assumption of normality dots are residuals. better than making a histogram

-    bottom left - assumption of equal variance - y axis as indicator for variance

-   bottom right - evaluates leverage  - important points in swaying the gradient and outliers

read stats textbooks apparently

no patterns means all good?

smooth.colour=NA - suppresses wiggly line

## 5.5.5 Now the interpretation 

can we reject the null hypothesis? Time to run the analysis - anova() and summary()

anova() doesn't perform an anova, it produces a table of statistics - sum-of-squares table.

summary() produces of the estimates of the coefficients of the line that is 'the model'

```{r}
anova(model_pgr)
```

Large F-values indicates a small error variance. With a single df, means a small p-value.

```{r}
summary(model_pgr)
```

## 5.5.6 From stats back to the figure 

final step in the workflow is to translate the model back to a figure.

```{r}
ggplot(plant_growth, aes(x=soil.moisture.content, y=plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  ylab("Plant Growth Rate (mm/week)") + 
  theme_bw()
```

geom_smooth(method = "lm") "is ggplot2-speak for 'shove a linear-model fitted line, and the standard error of the fit, using a flash transparent grey, onto my graph'"

# NEXT TIME: ANOVA
