---
title: "Homework 25"
author: "Evelyn Albrecht"
format: html
editor: visual
---

# 5.6 Analysis of variance: the one-way ANOVA

final example for this chapter (5) is the one-way ANOVA.

Explanatory variable is categorical variable

adding more levels than what we had in the compensation data.

## 5.6.1 Getting and plotting the data 

We will be looking at water flea growth rates. Do parasites (3 types) alter growth rates? We are adding a control treatment to the mix.

We are going to use the same workflow.

Lets do the basics

```{r}
rm(list=ls())
library(tidyverse)
library("here")
library(ggfortify)
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
glimpse(daphnia)
```

This data frame has 3 variables. we want to use two of them for the analysis.

we will be making a box-and-whisker plot to view variation in responses with a function of a categorical variable.

```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate)) + 
  geom_boxplot() + 
  theme_bw()
```

Coordinate flipping. We will flip the axes. This can help use visualize categorical data better.

```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate)) + 
  geom_boxplot() + 
  theme_bw() + 
  coord_flip()
```

This allows us to actually be able to read the names of the categories (they were all bunched on top of each other before).

From this it is noticeable that there is differences between the growth rates between the different treatments. The control seemingly produces the higher growth rate.

we expect the growth rate to be

1.  parasites overall bring growth rate down
2.  the extent to which they affect are P. ramosa \< M. bicuspidata \< P. perplexa.

we can also go on to estimate the average growth rate for each treatment and the degrees of freedom.

Time to construct a linear model the function is ...lm()

## 5.6.2 Construct the ANOVA

We will create model in the same way we did for the regression.

```{r}
model_grow <- lm(growth.rate ~ parasite, data=daphnia)
```

## 5.6.3 Check the Assumptions 

Time to evaluate the assumptions of this model

once again we are going to evaluate the assumptions in the same way: using autoplot

```{r}
autoplot(model_grow, smooth.colour = NA)
```

these figures suggest that everything is fine.

I really need to read a stats textbook

time to apply the anova() function to a ANOVA

## 5.6.4 Making an inference from a one-way ANOVA

First we are determining whether there is an effect at all of our treatments

```{r}
anova(model_grow)
```

The null hypothesis - all the groups come from the population with the same mean. The F-value is how likely that is to be true. small p values allows to reject null hypothesis.

## 5.6.5 Treatment contrasts 

produce the information in a summary table. treatment contrasts do this.

contrasts express coefficients taken from statistical models. look at "?contr.treatments"

```{r}
summary(model_grow)
```

treatment group "control" is missing, but not really, it has been labeled '(Intercept)'

R likes to alphabetize.

Treatment contrasts report differences between the reference level (control?) and other levels.

You can also get the means for each treatment using dplyr and then calculate the contrasts.

```{r}
sumDat <- daphnia %>%
  group_by(parasite) %>%
  summarise(meanGR = mean(growth.rate))
```

subtract mean of parasite from control.

We were lucky because the control was the reference.

Now we are going to produce a figure.

I'm not sure how we are suppose to do so.

# 5.7 Wrapping up

## take home messages 

-   Have a consistent workflow for analyses

-   Always make a graph that answers your question before touching any statistical tests.

-   interpret as much as you can, for example, slopes, intercepts, contrasts, and degrees of freedom, from the graph before doing statistics

-   check if your data violate any assumptions, before interpreting your model results

-   make a beautiful graph to communicate your results, and write a biologically focused sentence describing your results

-   R makes this all very easy
