---
title: "Homework 16"
author: "Evelyn Albrecht"
format: html
editor: visual
---

Setting up - clear R's environment and open packages

```{r}
rm(list=ls())
library(tidyverse)
library("here")
```

Create a new data frame using the donnerville data

```{r}
cover <- read.csv(here("data","donnerville_percent_cover.csv"))
```

Take peak of data in two ways I will be using the glimpse and structure fuctions to do so

```{r}
glimpse(cover)
str(cover)
```

### convert Site, Plot.type and cover_type into factor variables 

```{r}
cover$Site <-as.factor(cover$Site)
cover$Plot.type <- as.factor(cover$Plot.type)
cover$cover_type <- as.factor(cover$cover_type)
```

## transform date into date variable 
I'm doing so by reassigning the value of Date to consist of only numbers and then I will use the dmy() function to convert it's type to date

```{r}
fix_date <-which(cover$Date == "13-Sep-17") #make new data frame with all of the date listing of 13-Sep-17. Coincidentally, this is all of them
cover$Date[fix_date] <- "9-13-24" #override that character to be in an all number format
cover$Date <- mdy(cover$Date) #indicate to r that the "character" is in a date format 
```

### Check the data for any errors

I found a few discrepancies in the listing of the same level. I will now correct these using the which() function.

correcting sites
```{r}
levels(cover$Site)
fix_site1 <- which(cover$Site %in% c("donnerville 1","DONNERVILLE - 1", "Donnerville 1")) #making a data frame for rows that have these values 
cover$Site[fix_site1] <- "Donnerville 1" #choosing one of the values to be a uniform name for all of them
fix_site2 <- which(cover$Site %in% c("donnerville 2", "Donnerville 2"))
cover$Site[fix_site2] <- "Donnerville 2"
fix_site3 <- which(cover$Site %in% c("donnerville-3","Donnerville 3"))
cover$Site[fix_site3] <- "Donnerville 3"
cover$Site <- factor(cover$Site) #remind R that there are now 3 levels by doing this 
levels(cover$Site)
```
now doing the same thing for cover type
```{r}
levels(cover$cover_type)
fix_cover1 <- which(cover$cover_type %in% c("bare","Bare"))
cover$cover_type[fix_cover1] <- "Bare"
fix_cover2 <- which(cover$cover_type %in% c( "bryophtes","Bryophytes"))
cover$cover_type[fix_cover2] <- "Bryophytes"
fix_cover3 <- which(cover$cover_type %in% c("vasc ","Vascular"))
cover$cover_type[fix_cover3] <- "Vascular"
cover$cover_type <- factor(cover$cover_type)
levels(cover$cover_type)
```

## Determining the average percent cover by cover type for Donnerville 1
```{r}
cover %>% #within the cover data frame
  group_by(cover_type) %>% #sorting out means of     percent by the cover type
  summarise(meanPercent=mean(Percent)) #show me      the means
```
## now taking the average across all of the cover types 
```{r}
cover %>% 
  summarise(meanPercent=mean(Percent))
```

